Macroscopic Salt Rejection through Electrostatically Gated Porous Graphene

### 2019-01-14 paper/paper.tex
Correct typos

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Main text
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\clearpage{}
\linenumbers
\section{Introduction}
\label{sec:intro}
A central objective in the emerging field of nanofluidics is to
control transport phenomena at nanometer scale to enable and study
novel devices and systems. The reduction of characteristic dimensions
in nanofluidics gives rise to substantially different transport
properties in comparison to their bulk counterparts
\cite{Schoch_2008}. In nanopores, for example, the effect of
surface-mediated transport becomes dominant when the pore size is
comparable to the characteristic length of diffusion, enabling new
applications such as ionic diodes \cite{Karnik_2007}, field-effect
transistors \cite{Nam_2009}, desalination \cite{Heiranian_2015}, and
nanopore-based DNA sequencing \cite{Heerema_2016,Garaj_2013}. Porous,
atomically thin two-dimensional (2D) materials, such as graphene, are
emerging as ideal membrane materials due to their ultimate permeation
\cite{Suk_2010,Jiang_2009,Celebi_2014,Koenig_2012,Drahushuk_2012}. Early
research has suggested that using nanoporous graphene as a
desalination membrane enables excellent salt rejection by preventing
the passage of ions, maintaining a high water permeability
simultaneously.
\cite{Cohen_Tanugi_2012,Suk_2014,Cohen_Tanugi_2014,Cohen_Tanugi_2015,O_Hern_2014,O_Hern_2015,Surwade_2015,Walker_2017,Ghosh_2018}. Nevertheless,
the fabrication of subnanometer porous 2D membranes over a large area
with atomic precision remains technically challenging
\cite{Suk_2014,Rollings_2016,O_Hern_2012,Wang_2017}. Moreover, a small
portion of excessively large pores can significantly diminish salt
rejection.  Very recently, a number of reports have explored the
effect of surface charges on the selectivity of ion transport through
graphene nanopores, as a consequence of the electrostatic interactions
between solvated ions and the charged functional groups attached on
graphene, indicating that even large pores ($>$ 30nm) can be ion
selective \cite{Rollings_2016,Surwade_2014}. However, it remains
controversial if surface charges effectively allow controlled salt
-transport or -rejection, as the membrane charge and potential cannot
be precisely quantified and are sensitive to the surroundings and
chemical treatment \cite{Li_2008}. Therefore, there is a lack of
fundamental understanding of the interactions between the atomically
thin 2D membrane, its surface charge and the electrolyte
solution. Here, we study macroscopic salt rectification through a
sheet of large-area porous graphene using electrostatic gating,
thereby deepening the understanding of ion-graphene interaction and
sheding light on applications of nanoporous graphene for tailored ion
transport.

@@

Next, we carried out the same experiments using the PG-covered PCTE
membrane (PG-PCTE), without electrostatic gating. The measured
diffusive flux values, $\boldsymbol{J}_{\mathrm{PG0}}$ (green bars in
\Fig{} \ref{fig:2}b), remain at the same order of magnitude compared
with those of bare PCTE membrane, suggesting that the transport
resistance of PCTE $R_{\mathrm{PCTE}}$, is dominant over that of PG
without gating, $R_{\mathrm{G}}^{0}$. Using a simple circuit model, we
estimate the ratio $\delta$ between $R_{\mathrm{G}}^{0}$ and
$R_{\mathrm{PCTE}}$, such that
$\delta = R_{\mathrm{PCTE}}/R_{\mathrm{G}}^{0}$, to be $\sim$2.2,
which is in good agreement with the value derived from the
experimental data in \Fig{} \ref{fig:2}b (details see Supplementary
Section \ref{SI-sec:R-model}).  \todo[inline]{Make sure that the
  values used are correct.}

@@

We next discuss the effects of electrostatic gating on graphene. In
order to avoid any interference from the conductivity probe to the
membrane potential, the “three-interval” method, which has been used
in controlling the membrane potential in mesoporous carbon membranes
\cite{Surwade_2014}, is adopted here. \Fig{} \ref{fig:2}a
illustratively represents how the measured conductivity evolves with
time. Specifically, in the first interval, the gate voltage source is
turned off and the conductivity probe is on, allowing us to obtain the
conductivity-time profile with an average slope $s_{1}$ that
corresponds to the diffusive flux across the PG-PCTE membrane. In the
second interval, a voltage $V_{\mathrm{G}}$ is applied to the membrane
after turning off the conductivity probe, followed by the third
interval $s_{3}$, in which the conductivity probe is turned on again
to give the corresponding slope. A control experiment was performed to
ensure the copper tape was not in direct contact with the ionic
solution (Supplementary Section \ref{SI-sec:copper}). The slope
corresponding to the second interval $s_{2}$, as the conductivity
probe is off, is determined by linear interpolation of the
conductivities from the endpoint of interval 1 to the starting point
of interval 3. A decrease of slope observed in interval 2 indicates
that the total flux through PG-PCTE is reduced upon gating (i.e. salt
penetration through the PG pores is hindered). We define such
experimentally measured reduction rate as $\eta$, which is equivalent
to the decrease of total diffusive flux through PG-PCTE as follows:
\begin{equation}
  \label{eq:rejection}
  \eta = \frac{\bar{s} - s_{2}}{\bar{s}} = \frac{\boldsymbol{J}_{\mathrm{PG0}}
    - \boldsymbol{J}_{\mathrm{PG}}(V_{\mathrm{G}})}{\boldsymbol{J}_{\mathrm{PG0}}}
\end{equation}
where $\bar{s}$ is the average slope of $s_{1}$ and $s_{3}$, namely
$(s_{1} + s_{3})/2$, and
$\boldsymbol{J}_{\mathrm{PG}}(V_{\mathrm{G}})$ is the diffusive flux
across the PTCE-supported PG double layer as a function of
$V_{\mathrm{G}}$. PG and PCTE membranes effectively form a transport
system with series resistance $R_{\mathrm{PG}}$ (tunable by
$V_{\mathrm{G}}$) and $R_{\mathrm{PCTE}}$ (independent of
$V_{\mathrm{G}}$), respectively. Due to the dominance of
$R_{\mathrm{PCTE}}$ in the ionic transport, significant reduction of
total ionic flux can only be achieved, when
$R_{\mathrm{\mathrm{G}}}(V_{\mathrm{G}} \neq 0) \gg
R_{\mathrm{\mathrm{G}}}^{0}$. Using the circuit analog model, we can
extract the effective salt rejection ratio of PG $\xi$, defined as
$1 - R_{\mathrm{G}}^{0}/R_{\mathrm{G}}$ from the observable flux reduction ratio $\eta$ (see Supplementary
\Fig{} \ref{SI-fig:eta-all}), as a function of $V_{\mathrm{G}}$ as
(details see Supplementary Section \ref{SI-sec:R-model}):
\begin{equation}
\label{eq:xi-def}
\xi = \frac{(\delta+1) \eta}{\delta \eta + 1}
\end{equation}
The factor $\xi$ is a reliable and stable measure of salt rejection
through the PG membrane upon gating, as the salt species and initial
conditions may induce a degree of measurement uncertainty between
different samples.

@@

In order to quantitatively understand the observed $V_{\mathrm{G}}$
dependence, we develop a theory to model the coupling of graphene’s
elementary transport properties and EDL, in order to quantify the
ionic transport through a graphene nanopore. Under the assumption that
the time scale for the bulk concentration change is significantly
longer than that of ionic diffusion, the pseudo-steady state
approximation holds. Accordingly, the steady-state Nernst-Planck
equation describing the ionic transport in an electrolyte solution is
given by \cite{MacGillivray_1968}:
\begin{equation}
  \label{eq:pnp}
  \nabla \cdot \boldsymbol{J}_{i} = -\nabla \cdot (\frac{D_{i}}{k_{\mathrm{B}}T} c_{i} N_{\mathrm{A}} \nabla \mu_{i}) = 0
\end{equation}
where $\boldsymbol{J}$ is the mass flux, subscript \textit{i}
corresponds to the ionic species (anion or cation), $D$ is the diffusivity,
$k_{\mathrm{B}}$ is the Boltzmann constant, $T$ is temperature, $c$ is
the molar concentration, $N_{\mathrm{A}}$ is the Avogadro constant,
and $\mu_{i}$ is the electrochemical potential. Under the assumption
of ideal solution, it follows \cite{Kilic_2007}:
\begin{equation}
  \label{eq:mu}
  \nabla \mu_{i} = k_{\mathrm{B}} T \nabla \ln x_{i} + z_{i} e \nabla \psi
\end{equation}
 $x$ is the molar fraction in solution, $z$ is the ionic valence,
$e$ is the unit charge, and $\psi$ is the electric potential. On the
other hand, the Poisson equation describing the electric potential
distribution is given by:

@@

where $\hbar$ is the reduced Planck constant, and
$v_{\mathrm{F}}=1.1\times10^{6}$ m$\cdot$s$^{-1}$ is the Fermi
velocity of graphene (more details find Supplementary Section
\ref{SI-sec:charge-dens}).  Note that the PG film fabricated
experimentally is double-layer turbostratic (randomly aligned)
graphene, in which we assume the
$\Delta \phi_{\mathrm{G}} - \sigma_{\mathrm{G}}$ dependence follows
Eq. \eqref{eq:delta-phiG}. The theory proposed here was solved
self-consistently by discretizing
Eqs. \eqref{eq:pnp}-\eqref{eq:poisson} using the finite-element method
(FEM), in which the graphene surface potential is coupled with
Eqs. \eqref{eq:electro-neutral}-\eqref{eq:delta-phiG} that were solved
simultaneously to reach convergent numerical solutions of
$\sigma_{\mathrm{G}}$ and $\psi_{\mathrm{G}}$ for a given
$V_{\mathrm{G}}$. Clearly, the above model yields symmetric
characteristics for $\sigma_{\mathrm{G}}$ and $\psi_{\mathrm{G}}$ with
respect to $V_{\mathrm{G}}$ due to the symmetric band structure of
graphene (Eq. \eqref{eq:delta-phiG}). However, for graphene
synthesized by chemical vapor deposition (CVD) used in our
experiments, it is well-recognized that the electron traps are
inherently introduced during synthesis and patterning
\cite{Dean_2010}, effectively reducing the charge density and surface
potential on graphene $V_{\mathrm{G}}<0$ (Supplementary \Fig{}
\ref{SI-fig:trap}), which is related to the fact that observed $\xi$
is much lower within the $V_{\mathrm{G}}<0$ regime in \Fig{}
\ref{fig:2}c. With the nonideality in mind, hereafter, we compare the
calculations and experiments in the regime of $V_{\mathrm{G}}>0$. For
example, \Fig{} \ref{fig:3}c presents the calculated
$\sigma_{\mathrm{G}}$ and $\psi_{\mathrm{G}}$ as a function of
$V_{\mathrm{G}}$ considering the ionic diffusion through a single 20
nm-diameter nanopore drilled on a sheet of semi-infinite double layer
graphene that separates HCR containing KCl solution at $c_{0}$ = 0.1
mM, and LCR at $c_{0}/10$ in axisymmetric coordinates (details see
Supplementary Section \ref{SI-fig:scheme}). Indeed, the interplay between graphene
quantum capacitance and the EDL significantly reduces the graphene
surface potential $\psi_{\mathrm{G}}$ to $\sim$0.3 V at
$V_{\mathrm{G}}$ = 1.25 V, corresponding a surface charge density
$\sigma_{\mathrm{G}}$ of $\sim$0.08 C$\cdot$m$^{-2}$. Note that as we
mentioned earlier, an important merit of the system here is that the
surface charge density can be precisely determined and controlled,
rather than being treated as a fitting parameter as in most of the 2D
nanopore studies \cite{Rollings_2016}.

@@

We further reveal the cation and anion transport pathways by looking
into their electrochemical potentials (\Fig{} \ref{fig:4}b). By
increasing $V_{\mathrm{G}}$, a more positive $\psi_{\mathrm{G}}$ on
graphene surface reduces the cation concentration at the pore edge due
to the electrostatic interactions, thereby increasing its
concentration gradient at the pore center. On the other hand, the
anion concentration near the pore edge increases exponentially, such
that the concentration gradient at the pore center becomes small (see
Supplementary \Fig{} \ref{SI-fig:conc}). We also observe that
$\Delta \mu_{+}$ is dominated by the concentration, while
$\Delta \mu_{-}$ is much less than $\Delta \mu_{+}$ due to the balance
between the diffusion and drift potentials (see Supplementary \Fig{}
\ref{SI-fig:potential}). The observations confirm the distinct ionic
transport pathways upon applying a positive $V_{\mathrm{G}}$: pore
center for cations and pore edge for anions. \Fig{} \ref{fig:4}c
presents the calculated \textit{z}-component of the cation and anion
fluxes through the pore at the graphene plane (z = 0),
$\boldsymbol{J}_{z}$ ,as a function of the normalized radius,
$r/r_{\mathrm{G}}$ , where $r$ is the radial coordinate and
$r_{\mathrm{G}}$ is the pore radius, at different $V_{\mathrm{G}}$
values. Note that $\boldsymbol{J}_{z}$ is negative because the LCR is
placed at z$<$0 in the simulation box. At $V_{\mathrm{G}}$ = 0 (pure
diffusion), the anion and cation flux profiles are identical, with a
higher flux near the pore edge, which is expected, as the
concentration gradient is higher. By gradually increasing
$V_{\mathrm{G}}$, both fluxes are reduced throughout the pore, while
the degree of reduction is more pronounced at the pore edge for the
cation, and at the pore center for the anion, following the scenarios
we discussed above. Accordingly, the integrated fluxes across the
pore, $|\boldsymbol{J}_{z}|$, are reduced with $V_{\mathrm{G}}$
(\Fig{} \ref{fig:4}d). A total flux reduction of up to 80\% is
predicted. Another interesting finding here is that, by increasing
$V_{\mathrm{G}}$, the nanopore transport preferentially allows cations
over anions, or in other words, the anion flux is more reduced with
$V_{\mathrm{G}}$ (see \Fig{} \ref{fig:4}c), known as the ion
selectivity of graphene nanopore \cite{Rollings_2016}. In our setup,
the imbalance between cation and anion fluxes near the graphene
nanopore is compensated by the EDL formation at the counter-electrode
in LCR, which maintains the electroneutrality of the system. Notably,
since atomically thin graphene samples are used, the mechanism of salt
rejection is clearly distinct from that in capacitive deionization
\cite{Biesheuvel_2010}, where electrodes with large surface area are
required.

The above mechanistic findings, nevertheless, do not fully clarify the
experimentally observed reduction of diffusive flux upon
gating. Indeed, the same physical mechanism also governs the ionic
transport through an IFET, in which a nonzero electric potential at
the nanopore center usually results in an increase of ionic
conductivity \cite{Nam_2009,Lee_2015,Feng_2016}. To this end, we
further increased the graphene surface potential $\psi_{\mathrm{G}}$
considering the same system (details see Supplementary Section
\ref{SI-sec:simu-res}). Intriguingly, by increasing $\psi_{\mathrm{G}}$
larger than 400 mV, the ionic flux starts to increase, reversing the
trend observed at the low $\psi_{\mathrm{G}}$ regime, due to (i) the
drift flux become dominant over diffusion flux and (ii) the
concentration near the pore edge (prefered path) greatly increases
(details see Supplementary Section \ref{SI-sec:simu-res}, Supplementary
Figs. \ref{SI-fig:reverse} and \ref{SI-fig:large-V}). This level of
$\psi_{\mathrm{G}}$ cannot be reached experimentally in graphene
membrane, as this requires a $V_{\mathrm{G}}$ larger than 2.0 V, which
triggers the electrochemical reactions \cite{Toh_2011}. On the other
hand, in an IFET, the electric potential on the pore wall is
considerably higher, equivalent to $V_{\mathrm{G}}$ due to the
metallic nature of gate electrode (having an infinitely-large quantum
capacitance). We therefore conclude that the quantum
capacitance-induced nonlinear damping in the surface potential results
in the observed salt rejection.  Following the above discussions, we
further investigate the physical limits for the biased graphene
nanopores. Specifically, the salt rejection characteristics are
controlled by: (i) the overlap of EDL inside the nanopore, essentially
controlled by two length scales of $\lambda_{\mathrm{D}}$ and
$r_{\mathrm{G}}$, and (ii) the graphene surface potential
$\psi_{\mathrm{G}}$ controlled by $V_{\mathrm{G}}$ following
Eq. \eqref{eq:Vg}. To this end, we calculate $\psi_{\mathrm{G}}$ as a
function of $\lambda_{\mathrm{D}} / r_{\mathrm{G}}$ and
$V_{\mathrm{G}}$ (\Fig{} \ref{fig:5}a). Clearly, when the bulk
concentration $c_{0}$ increases, a higher $V_{\mathrm{G}}$ is required
to reach the same level of $\psi_{\mathrm{G}}$. Consequently, as
revealed in \Fig{} \ref{fig:5}b, the calculated salt rejection factor
$\xi$ increases with both $\lambda_{\mathrm{D}} / r_{\mathrm{G}}$ and
$V_{\mathrm{G}}$, in line with the experimentally observed
$\xi - c_{0}$ dependence. We also find the overall trends for
$\psi_{\mathrm{G}}$ and $\xi$ with respect to
$\lambda_{\mathrm{D}}/r_{\mathrm{G}}$ and $V_{\mathrm{G}}$ are very
similar. More importantly, based on our theoretical prediction, with
$V_{\mathrm{G}}$ = 1.25 V and $\lambda_{\mathrm{D}} / r_{\mathrm{G}}$
= 1.0, a high value of $\xi$ of $>$ 0.9 can be achieved.

% As discussed previously, the diffusive transport resistance of ungated patterned graphene 
% is much smaller than that of the PCTE. When gating the graphene membrane,
% the overall flux, however, reduces to $\sim$ 55 $\%$
% of the ungated flux. This indicates, that R$_{graphene}$ $\approx$ R$_{PCTE}$, 
% a 26 fold increase of the resistance with respect to the ungated case. 
% This finding indicates, that the experimentally observed $\xi$ for graphene alone 
% is indeed higher than the 55$\%$, which is observed for the PG-PCTE combined.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Main text
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\clearpage{}
\linenumbers
\section{Introduction}
\label{sec:intro}
A central objective in the emerging field of nanofluidics is to
control transport phenomena at nanometer scale to enable and study
novel devices and systems. The reduction of characteristic dimensions
in nanofluidics gives rise to substantially different transport
properties in comparison to their bulk counterparts
\cite{Schoch_2008}. In nanopores, for example, the effect of
surface-mediated transport becomes dominant when the pore size is
comparable to the characteristic length of diffusion, enabling new
applications such as ionic diodes \cite{Karnik_2007}, field-effect
transistors \cite{Nam_2009}, desalination \cite{Heiranian_2015}, and
nanopore-based DNA sequencing \cite{Heerema_2016,Garaj_2013}. Porous,
atomically thin two-dimensional (2D) materials, such as graphene, are
emerging as ideal membrane materials due to their ultimate permeation
\cite{Suk_2010,Jiang_2009,Celebi_2014,Koenig_2012,Drahushuk_2012}. Early
research has suggested that using nanoporous graphene as a
desalination membrane enables excellent salt rejection by preventing
the passage of ions, maintaining a high water permeability
simultaneously.
\cite{Cohen_Tanugi_2012,Suk_2014,Cohen_Tanugi_2014,Cohen_Tanugi_2015,O_Hern_2014,O_Hern_2015,Surwade_2015,Walker_2017,Ghosh_2018}. Nevertheless,
the fabrication of subnanometer porous 2D membranes over a large area
with atomic precision remains technically challenging
\cite{Suk_2014,Rollings_2016,O_Hern_2012,Wang_2017}. Moreover, a small
portion of excessively large pores can significantly diminish salt
rejection.  Very recently, a number of reports have explored the
effect of surface charges on the selectivity of ion transport through
graphene nanopores, as a consequence of the electrostatic interactions
between solvated ions and the charged functional groups attached on
graphene, indicating that even large pores ($>$ 30nm) can be ion
selective \cite{Rollings_2016,Surwade_2014}. However, it remains
controversial if surface charges effectively allow controlled
salt-transport or -rejection, as the membrane charge and potential
cannot be precisely quantified and are sensitive to the surroundings
and chemical treatment \cite{Li_2008}. Therefore, there is a lack of
fundamental understanding of the interactions between the atomically
thin 2D membrane, its surface charge and the electrolyte
solution. Here, we study macroscopic salt rectification through a
sheet of large-area porous graphene using electrostatic gating,
thereby deepening the understanding of ion-graphene interaction and
shedding light on applications of nanoporous graphene for tailored ion
transport.

@@

Next, we carried out the same experiments using the PG-covered PCTE
membrane (PG-PCTE), without electrostatic gating. The measured
diffusive flux values, $\boldsymbol{J}_{\mathrm{PG0}}$ (green bars in
\Fig{} \ref{fig:2}b), remain at the same order of magnitude compared
with those of bare PCTE membrane, suggesting that the transport
resistance of PCTE $R_{\mathrm{PCTE}}$, is dominant over that of PG
without gating, $R_{\mathrm{G}}^{0}$. Using a simple circuit model, we
estimate the ratio $\delta$ between $R_{\mathrm{G}}^{0}$ and
$R_{\mathrm{PCTE}}$, such that
$\delta = R_{\mathrm{PCTE}}/R_{\mathrm{G}}^{0}$, to be $\sim$2.2,
which is in good agreement with the value derived from the
experimental data in \Fig{} \ref{fig:2}b (details see Supplementary
Section \ref{SI-sec:R-model}).
% \todo[inline]{Make sure that the
  % values used are correct.}

@@

We next discuss the effects of electrostatic gating on graphene. In
order to avoid any interference from the conductivity probe to the
membrane potential, the ``three-interval'' method, which has been used
in controlling the membrane potential in mesoporous carbon membranes
\cite{Surwade_2014}, is adopted here. \Fig{} \ref{fig:2}a
illustratively represents how the measured conductivity evolves with
time. Specifically, in the first interval, the gate voltage source is
turned off and the conductivity probe is on, allowing us to obtain the
conductivity-time profile with an average slope $s_{1}$ that
corresponds to the diffusive flux across the PG-PCTE membrane. In the
second interval, a voltage $V_{\mathrm{G}}$ is applied to the membrane
after turning off the conductivity probe, followed by the third
interval $s_{3}$, in which the conductivity probe is turned on again
to give the corresponding slope. A control experiment was performed to
ensure the copper tape was not in direct contact with the ionic
solution (Supplementary Section \ref{SI-sec:copper}). The slope
corresponding to the second interval $s_{2}$, as the conductivity
probe is off, is determined by linear interpolation of the
conductivities from the endpoint of interval 1 to the starting point
of interval 3. A decrease of slope observed in interval 2 indicates
that the total flux through PG-PCTE is reduced upon gating (i.e. salt
penetration through the PG pores is hindered). We define such
experimentally measured reduction rate as $\eta$, which is equivalent
to the decrease of total diffusive flux through PG-PCTE as follows:
\begin{equation}
  \label{eq:rejection}
  \eta = \frac{\bar{s} - s_{2}}{\bar{s}} = \frac{\boldsymbol{J}_{\mathrm{PG0}}
    - \boldsymbol{J}_{\mathrm{PG}}(V_{\mathrm{G}})}{\boldsymbol{J}_{\mathrm{PG0}}}
\end{equation}
where $\bar{s}$ is the average slope of $s_{1}$ and $s_{3}$, namely
$(s_{1} + s_{3})/2$, and
$\boldsymbol{J}_{\mathrm{PG}}(V_{\mathrm{G}})$ is the diffusive flux
across the PTCE-supported PG double layer as a function of
$V_{\mathrm{G}}$. PG and PCTE membranes effectively form a transport
system with series resistance $R_{\mathrm{PG}}$ (tunable by
$V_{\mathrm{G}}$) and $R_{\mathrm{PCTE}}$ (independent of
$V_{\mathrm{G}}$), respectively. Due to the dominance of
$R_{\mathrm{PCTE}}$ in the ionic transport, significant reduction of
total ionic flux can only be achieved, when
$R_{\mathrm{\mathrm{G}}}(V_{\mathrm{G}} \neq 0) \gg
R_{\mathrm{\mathrm{G}}}^{0}$. Using the circuit analog model, we can
extract the effective salt rejection ratio of PG $\xi$, (defined as
$1 - R_{\mathrm{G}}^{0}/R_{\mathrm{G}}$) from the observable flux
reduction ratio $\eta$ (see Supplementary \Fig{}
\ref{SI-fig:eta-all}), as a function of $V_{\mathrm{G}}$ as (details
see Supplementary Section \ref{SI-sec:R-model}):
\begin{equation}
\label{eq:xi-def}
\xi = \frac{(\delta+1) \eta}{\delta \eta + 1}
\end{equation}
The factor $\xi$ is a reliable and stable measure of salt rejection
through the PG membrane upon gating, as the salt species and initial
conditions may induce a degree of measurement uncertainty between
different samples.

@@

In order to quantitatively understand the observed $V_{\mathrm{G}}$
dependence, we develop a theory to model the coupling of graphene's
elementary transport properties and EDL, in order to quantify the
ionic transport through a graphene nanopore. Under the assumption that
the time scale for the bulk concentration change is significantly
longer than that of ionic diffusion, the pseudo-steady state
approximation holds. Accordingly, the steady-state Nernst-Planck
equation describing the ionic transport in an electrolyte solution is
given by \cite{MacGillivray_1968}:
\begin{equation}
  \label{eq:pnp}
  \nabla \cdot \boldsymbol{J}_{i} = -\nabla \cdot (\frac{D_{i}}{k_{\mathrm{B}}T} c_{i} N_{\mathrm{A}} \nabla \mu_{i}) = 0
\end{equation}
where $\boldsymbol{J}$ is the mass flux, subscript \textit{i}
corresponds to the ionic species (anion or cation), $D$ is the diffusivity,
$k_{\mathrm{B}}$ is the Boltzmann constant, $T$ is temperature, $c$ is
the molar concentration, $N_{\mathrm{A}}$ is the Avogadro constant,
and $\mu_{i}$ is the electrochemical potential. Under the assumption
of ideal solution, it follows \cite{Kilic_2007}:
\begin{equation}
  \label{eq:mu}
  \nabla \mu_{i} = k_{\mathrm{B}} T \nabla \ln x_{i} + z_{i} e \nabla \psi
\end{equation}
 $x$ is the molar fraction in solution, $z$ is the ionic valence,
$e$ is the unit charge, and $\psi$ is the electric potential. On the
other hand, the Poisson equation describing the electric potential
distribution is given by:

@@

where $\hbar$ is the reduced Planck constant, and
$v_{\mathrm{F}}=1.1\times10^{6}$ m$\cdot$s$^{-1}$ is the Fermi
velocity of graphene (more details find Supplementary Section
\ref{SI-sec:charge-dens}).  Note that the PG film fabricated
experimentally is double-layer turbostratic (randomly aligned)
graphene, in which we assume the
$\Delta \phi_{\mathrm{G}} - \sigma_{\mathrm{G}}$ dependence follows
Eq. \eqref{eq:delta-phiG}. The theory proposed here was solved
self-consistently by discretizing
Eqs. \eqref{eq:pnp}-\eqref{eq:poisson} using the finite-element method
(FEM), in which the graphene surface potential is coupled with
Eqs. \eqref{eq:electro-neutral}-\eqref{eq:delta-phiG} that were solved
simultaneously to reach convergent numerical solutions of
$\sigma_{\mathrm{G}}$ and $\psi_{\mathrm{G}}$ for a given
$V_{\mathrm{G}}$. Clearly, the above model yields symmetric
characteristics for $\sigma_{\mathrm{G}}$ and $\psi_{\mathrm{G}}$ with
respect to $V_{\mathrm{G}}$ due to the symmetric band structure of
graphene (Eq. \eqref{eq:delta-phiG}). However, for graphene
synthesized by chemical vapor deposition (CVD) used in our
experiments, it is well-recognized that the electron traps are
inherently introduced during synthesis and patterning
\cite{Dean_2010}, effectively reducing the charge density and surface
potential on graphene $V_{\mathrm{G}}<0$ (Supplementary \Fig{}
\ref{SI-fig:trap}), which is related to the fact that observed $\xi$
is much lower within the $V_{\mathrm{G}}<0$ regime in \Fig{}
\ref{fig:2}c. With the nonideality in mind, hereafter, we compare the
calculations and experiments in the regime of $V_{\mathrm{G}}>0$. For
example, \Fig{} \ref{fig:3}c presents the calculated
$\sigma_{\mathrm{G}}$ and $\psi_{\mathrm{G}}$ as a function of
$V_{\mathrm{G}}$ considering the ionic diffusion through a single 20
nm-diameter nanopore drilled on a sheet of semi-infinite double layer
graphene that separates HCR containing KCl solution at $c_{0}$ = 0.1
mM, and LCR at $c_{0}/10$ in axisymmetric coordinates (details see
Supplementary Section \ref{SI-fig:scheme}). Indeed, the interplay
between graphene quantum capacitance and the EDL significantly reduces
the graphene surface potential $\psi_{\mathrm{G}}$ to $\sim$0.3 V at
$V_{\mathrm{G}}$ = 1.25 V, corresponding a surface charge density
$\sigma_{\mathrm{G}}$ of $\sim$0.08 C$\cdot$m$^{-2}$. Note that as we
mentioned earlier, an important merit of the system here is that the
surface charge density can be precisely determined and controlled,
rather than being treated as a fitting parameter as in most of the 2D
nanopore studies \cite{Rollings_2016}. We anticipate that the
gate-controlled ionic transport can also be achieved by replacing
graphene with other types of conducting 2D materials, although the
gate voltage required to achieve the same order of salt rejection may
be higher than graphene due to larger quantum capacitances.

@@

We further reveal the cation and anion transport pathways by looking
into their electrochemical potentials (\Fig{} \ref{fig:4}b). By
increasing $V_{\mathrm{G}}$, a more positive $\psi_{\mathrm{G}}$ on
graphene surface reduces the cation concentration at the pore edge due
to the electrostatic interactions, thereby increasing its
concentration gradient at the pore center. On the other hand, the
anion concentration near the pore edge increases exponentially, such
that the concentration gradient at the pore center becomes small (see
Supplementary \Fig{} \ref{SI-fig:conc}). We also observe that
$\Delta \mu_{+}$ is dominated by the concentration, while
$\Delta \mu_{-}$ is much less than $\Delta \mu_{+}$ due to the balance
between the diffusion and drift potentials (see Supplementary \Fig{}
\ref{SI-fig:potential}). The observations confirm the distinct ionic
transport pathways upon applying a positive $V_{\mathrm{G}}$: pore
center for cations and pore edge for anions. \Fig{} \ref{fig:4}c
presents the calculated \textit{z}-component of the cation and anion
fluxes through the pore at the graphene plane (z = 0),
$\boldsymbol{J}_{z}$ ,as a function of the normalized radius,
$r/r_{\mathrm{G}}$ , where $r$ is the radial coordinate and
$r_{\mathrm{G}}$ is the pore radius, at different $V_{\mathrm{G}}$
values. Note that $\boldsymbol{J}_{z}$ is negative because the LCR is
placed at z$<$0 in the simulation box. At $V_{\mathrm{G}}$ = 0 (pure
diffusion), the anion and cation flux profiles are identical, with a
higher flux near the pore edge, which is expected, as the
concentration gradient is higher. By gradually increasing
$V_{\mathrm{G}}$, both fluxes are reduced throughout the pore, while
the degree of reduction is more pronounced at the pore edge for the
cation, and at the pore center for the anion, following the scenarios
we discussed above. Accordingly, the integrated fluxes across the
pore, $|\boldsymbol{J}_{z}|$, are reduced with $V_{\mathrm{G}}$
(\Fig{} \ref{fig:4}d). A total flux reduction of $\sim$90\% is
predicted. Another interesting finding here is that, by increasing
$V_{\mathrm{G}}$, the nanopore transport preferentially allows cations
over anions, or in other words, the anion flux is more reduced with
$V_{\mathrm{G}}$ (see \Fig{} \ref{fig:4}c), known as the ion
selectivity of graphene nanopore \cite{Rollings_2016}. In our setup,
the imbalance between cation and anion fluxes near the graphene
nanopore is compensated by the EDL formation at the counter-electrode
in LCR, which maintains the electroneutrality of the system. Notably,
since atomically thin graphene samples are used, the mechanism of salt
rejection is clearly distinct from that in capacitive deionization
\cite{Biesheuvel_2010}, where electrodes with large surface area are
required.

The above mechanistic findings, nevertheless, do not fully clarify the
experimentally observed reduction of diffusive flux upon
gating. Indeed, the same physical mechanism also governs the ionic
transport through an IFET, in which a nonzero electric potential at
the nanopore center usually results in an increase of ionic
conductivity \cite{Nam_2009,Lee_2015,Feng_2016}. To this end, we
further increased the graphene surface potential $\psi_{\mathrm{G}}$
considering the same system (details see Supplementary Section
\ref{SI-sec:simu-res}). Intriguingly, by increasing
$\psi_{\mathrm{G}}$ larger than 400 mV, the ionic flux starts to
increase, reversing the trend observed at the low $\psi_{\mathrm{G}}$
regime, due to (i) the drift flux become dominant over diffusion flux
and (ii) the concentration near the pore edge (preferred path) greatly
increases (details see Supplementary Section \ref{SI-sec:simu-res},
Supplementary Figs. \ref{SI-fig:reverse} and
\ref{SI-fig:large-V}). This level of $\psi_{\mathrm{G}}$ cannot be
reached experimentally in graphene membrane, as this requires a
$V_{\mathrm{G}}$ larger than 2.0 V, which triggers the electrochemical
reactions \cite{Toh_2011}. On the other hand, in an IFET, the electric
potential on the pore wall is considerably higher, equivalent to
$V_{\mathrm{G}}$ due to the metallic nature of gate electrode (having
an infinitely-large quantum capacitance). We therefore conclude that
the quantum capacitance-induced nonlinear damping in the surface
potential results in the observed salt rejection.  Following the above
discussions, we further investigate the physical limits for the biased
graphene nanopores. Specifically, the salt rejection characteristics
are controlled by: (i) the overlap of EDL inside the nanopore,
essentially controlled by two length scales of $\lambda_{\mathrm{D}}$
and $r_{\mathrm{G}}$, and (ii) the graphene surface potential
$\psi_{\mathrm{G}}$ controlled by $V_{\mathrm{G}}$ following
Eq. \eqref{eq:Vg}. To this end, we calculate $\psi_{\mathrm{G}}$ as a
function of $\lambda_{\mathrm{D}} / r_{\mathrm{G}}$ and
$V_{\mathrm{G}}$ (\Fig{} \ref{fig:5}a). Clearly, when the bulk
concentration $c_{0}$ increases, a higher $V_{\mathrm{G}}$ is required
to reach the same level of $\psi_{\mathrm{G}}$. Consequently, as
revealed in \Fig{} \ref{fig:5}b, the calculated salt rejection factor
$\xi$ increases with both $\lambda_{\mathrm{D}} / r_{\mathrm{G}}$ and
$V_{\mathrm{G}}$, in line with the experimentally observed
$\xi - c_{0}$ dependence. We also find the overall trends for
$\psi_{\mathrm{G}}$ and $\xi$ with respect to
$\lambda_{\mathrm{D}}/r_{\mathrm{G}}$ and $V_{\mathrm{G}}$ are very
similar. More importantly, based on our theoretical prediction, with
$V_{\mathrm{G}}$ = 1.25 V and $\lambda_{\mathrm{D}} / r_{\mathrm{G}}$
= 1.0, a high value of $\xi$ up to $\sim$0.95 can be achieved. Nevertheless
the salt rejection of the gated graphene platform may be further
improved after considering the following aspects: (i) the mean-field
theory used here may be incorrect to be applied to pores smaller than
5 nm, where individual particle properties of solvated ions can be
important \cite{Jain_2015}, (ii) the specific ion-graphene adsorption
may also be included to account for the pore selectivity between salt
species \cite{Rollings_2016}.
